---
title: "Reproducible Research: Peer Assessment 1"
author: "pitambar Gautam: template after RDPeng"
date: "2020/10/9"
output: html_document
---

## Loading and preprocessing the data
```{r data_load_chunk, echo=TRUE}
activity <- read.table (file ="activity.csv",header = TRUE, sep =",")
head(activity)
tail(activity)
summary(activity)
```
## Histogram of Daily Steps data
```{r histogram_chunk, echo=TRUE}
## install.packages("ggplot2")
require(ggplot2) 
daily_steps <- aggregate(steps ~ date, activity, sum)
head(daily_steps)
str(daily_steps)
hist_ds<- hist(daily_steps$steps, main = "Histogram of Daily Steps", xlab = "Daily Steps", ylab = "No. of Days")

```


```{r summary_of_ daily_steps_chunk, echo=TRUE}
## Summary of Daily Steps data
summary(daily_steps$steps)

```

## Average daily activity pattern:
## Mean and Median number of steps taken per day
```{r mean_median_chunk, echo=TRUE}
mean_daily_steps <- mean(daily_steps$steps)
mean_daily_steps
median_daily_steps <- median(daily_steps$steps)
median_daily_steps
```
## Time-series of total steps taken per day
```{r time_series_plot_chunk, echo=TRUE}
daily_steps$date <- as.Date(daily_steps$date)
daily_steps <- daily_steps[order(daily_steps$date) , ]
plot(daily_steps$steps~daily_steps$date, type="b", lwd=3, col=rgb(0.1,0.7,0.1,0.8), main = "Time-series Steps Data",ylab="Total Daily Steps", xlab="Date" , bty="l" , pch=20 , cex=4) + abline(h=seq(0,100,10), col="grey", lwd=0.8)
```
## 5-minute time interval total steps Time-series data
```{r interval_w_max_steps_chunk, echo=TRUE}
interval_steps <- aggregate(steps ~ interval, activity, mean)
interval_steps <- interval_steps[order(interval_steps$steps, decreasing = TRUE) , ]
head(interval_steps)
interval_maxsteps <-interval_steps$interval[1]
interval_maxsteps
```

```{r interval_mean_addcol_chunk, echo=TRUE}
activity["interval_meansteps"]<- NA
activity$interval_meansteps = interval_steps$steps
# activity <- activity[order(activity$interval_meansteps, decreasing = TRUE) , ]
head(activity)
```

## Imputing missing values (replace by interval mean)

```{r fill_steps_with-interval-meansteps_chunk, echo=TRUE}

# No. of missing values from sumary above = 2304
activity["filled_steps"]<- NA # an empty (na) column added
activity$filled_steps = activity$steps  # content replaced by steps column data
# code to replace by overall mean steps:
# activity$filled_steps = ifelse(is.na(activity$filled_steps),ave(activity$steps,  (+contd..)
# FUN = function (x) mean (x, na.rm =TRUE)), activity$filled_steps)
activity$filled_steps = ifelse(is.na(activity$filled_steps), activity$interval_meansteps, activity$filled_steps)
head(activity)
```

## Average daily activity pattern (after imputing data):

```{r histogram_chunk_for_filled_daily_steps_chunk, echo=TRUE}
## Histogram od Daily Steps (revised) data 
## install.packages("ggplot2")
require(ggplot2) 
daily_steps_filled <- aggregate(filled_steps ~ date, activity, sum)
head(daily_steps_filled)
str(daily_steps_filled)
hist_dsfilled<- hist(daily_steps_filled$filled_steps, main = "Histogram of Daily Steps (revised)", xlab = "Daily Steps (revised)", ylab = "No. of Days")
```
## Summary of Daily Steps Data:
```{r summary_of_ daily_steps_filled_chunk, echo=TRUE}
summary(daily_steps_filled$filled_steps)

```
## Mean and Median number of steps taken per day:
```{r mean_median_for_filled_steps_chunk, echo=TRUE}
mean_daily_steps_filled <- mean(daily_steps_filled$filled_steps)
mean_daily_steps_filled
median_daily_steps_filled <- median(daily_steps_filled$filled_steps)
median_daily_steps_filled
```
## Differences in activity patterns between weekdays and weekends:
```{r weekdays_weekends_discrimination_chunk, echo=TRUE}
activity$date <-as.Date(activity$date) #conversion to date format for conversion with weekdays to week's day
activity$day_class <-  ifelse(as.POSIXlt(activity$date)$wday %in% c(0,6), 'weekend', 'weekday')
head(activity)
```

```{r time_series_plots_for_weekdays_n_weekends_chunk, echo=TRUE}
interval_steps_filled <- aggregate(filled_steps ~ interval + day_class, data = activity, mean)
head(interval_steps_filled)
ggplot(interval_steps_filled, aes(interval, filled_steps)) + 
    geom_line() + 
    facet_grid(day_class ~ .) + xlab("interval (each 5 minutes)") + ylab("Average Number of Steps")
  
```
